---
title: "Honors Thesis Working Document"
author: "Boston Lee"
date: "9/19/2021"
output: 
  pdf_document:
    extra_dependencies: ["booktabs"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(knitr)
library(tidyverse)
```


## Numerical summaries

```{r read-data}
mental_health_df <- read_csv("../data/preprocessed/mental_health_2018.csv")
mental_health_df$substate_region_id <- as.character(mental_health_df$substate_region_id)
```



# Modeling


## General Modeling Functions

We can create a general framework for including variables in our model
for a given dataset (of course requiring that the predictor and response
variables are present). The model will need to be a logistic regression
model, given that all of the proposed responses are proportions.

```{r nonspatial-lm-funcs}
suicidal_thoughts_lm <- function(df) {
  suicidal_thoughts_lm <- lm(
    prop_suicidal_thoughts ~ .,
    data = select(
      df, -c(region_id, prop_alcohol_use_disorder, prop_MDE)
    )
  )
  return(suicidal_thoughts_lm)
}

alcohol_use_disorder_lm <- function(df) {
  alcohol_use_disorder_lm <- lm(
    prop_alcohol_use_disorder ~ .,
    data = select(
      df, -c(region_id, prop_suicidal_thoughts, prop_MDE)
    )
  )
  return(alcohol_use_disorder_lm)
}

MDE_lm <- function(df) {
  MDE_lm <- lm(
    prop_MDE ~ .,
    data = select(
      df, -c(region_id, prop_alcohol_use_disorder, prop_suicidal_thoughts)
    )
  )
  return(MDE_lm)
}
```

## Datasets for primary analysis

```{r base-data}
base_model_df <- mental_health_df %>%
  mutate(region_id = as.factor(paste(as.character(substate_region_id), as.character(state_fips), sep = "-"))) %>%
  group_by(region_id) %>%
  mutate(pop_weight = total_pop / sum(total_pop)) %>%
  summarize_if(is.numeric, ~ sum(. * pop_weight)) %>%
  select(-c(total_pop, pop_weight))
```

We will have three primary groups of models. Each group will consist of three models, with each of the SAMHSA variables as responses. The datasets will be as follows:


1. Predictor: Proportion below poverty level; control variables

```{r average-data}
primary_model_df <- base_model_df %>% select(-median_household_income)
print(primary_model_df)
```

1. Predictor: Proportion below poverty level; median income;  control variables

Here is the data as a population-weighted average, only including a unique row
ID and the numeric variables:

```{r average-data-income}
model_with_income_df <- base_model_df
```

1. Predictor: Proportion below poverty level; indicator for state; control variables

```{r average-data-state}
model_with_state_df <- mental_health_df %>%
  mutate(
    region_id = as.factor(paste(as.character(substate_region_id), as.factor(state_fips), sep = "-")),
    state = as.factor(state)
  ) %>%
  group_by(state, region_id) %>%
  mutate(pop_weight = total_pop / sum(total_pop)) %>%
  summarize_if(is.numeric, ~ sum(. * pop_weight)) %>%
  select(-c(total_pop, pop_weight, median_household_income))
print(model_with_state_df)
```

1. Predictor: Proportion below poverty level; indicator for state; *no* control variables

```{r data-state-no-adjust}
model_with_state_no_adjust_df <- model_with_state_df %>%
  select(prop_below_poverty_level, state, prop_MDE, prop_suicidal_thoughts, prop_alcohol_use_disorder, region_id)
print(model_with_state_no_adjust_df)
```


## Maps accurate to modeling data


## Maps

```{r map-level-model-df}
modified_full_df <- mental_health_df %>%
  mutate(
    region_id = as.factor(paste(as.character(substate_region_id), as.factor(state_fips), sep = "-"))
  ) %>%
  select(c(region_id, substate_region_id, county_fips, fips))
map_model_df <- modified_full_df %>%
  full_join(model_with_income_df)
# map_model_df <- model_df %>%
#   full_join(modified_full_df)
```

```{r map-func-county}
map_var <- function(mental_health_df, var_name, legend_title = "", percent_flag = FALSE) {
  # Grab county FIPS, pad with zeros to match data, and pull out columns
  maps::county.fips %>%
    as_tibble() %>%
    mutate(fips = str_pad(fips, 5, "left", "0")) %>%
    extract(polyname, c("region", "subregion"), "^([^,]+),([^,]+)$") ->
  fips_df
  head(fips_df)
  # Join county map objects with FIPS, and then use that linkage
  # to join the mental health data
  map_data("county") %>%
    left_join(fips_df) %>%
    left_join(mental_health_df) ->
  fips_county_df
  head(fips_county_df)
  # Plot overall
  fips_county_df %>%
    ggplot(aes(long, lat, group = group)) +
    geom_polygon(aes_string(fill = var_name), color = "gray70", size = 0.05) +
    coord_map() +
    # ggtitle(paste("Map of ", var_name, " across the United States")) +
    theme(plot.margin = grid::unit(c(0, 0, 0, 0), "mm")) +
    theme_void() +
    theme(legend.position = "bottom", legend.key.width = unit(2.5, "cm")) -> temp_map
  guide <- guide_colorbar(direction = "horizontal", title.position = "top")
  if (percent_flag) {
    return(temp_map +
      scale_fill_distiller(
        palette = "Spectral", labels = scales::percent,
        guide = guide, name = legend_title
      ))
  } else {
    return(temp_map +
      scale_fill_distiller(palette = "Spectral", guide = guide, name = legend_title))
  }
}

map_and_save_var <- function(filename, mental_health_df, var_name, legend_title = "", percent_flag = FALSE) {
  map_plot <- map_var(mental_health_df, var_name, legend_title, percent_flag)
  ggsave(filename, map_plot, width = 7, height = 5)
}
```

```{r plot-save-results-figs}
# Main results figures
map_and_save_var("./maps/prop_MDE_aggregated.png", map_model_df, "prop_MDE", "Percent of reported MDEs", percent_flag = TRUE)
map_and_save_var("./maps/prop_suicidal_thoughts_aggregated.png", map_model_df, "prop_suicidal_thoughts", "Percent of reported suicidal thoughts", percent_flag = TRUE)
map_and_save_var("./maps/prop_alcohol_use_disorder_aggregated.png", map_model_df, "prop_alcohol_use_disorder", "Percent of reported alcohol use disorder", percent_flag = TRUE)
map_and_save_var("./maps/median_household_income_aggregated.png", map_model_df, "median_household_income", "Median Household Income", percent_flag = FALSE)
map_and_save_var("./maps/prop_below_poverty_level_aggregated.png", map_model_df, "prop_below_poverty_level", "Percent of the population below poverty level", percent_flag = TRUE)
```


## Numerical Summaries

```{r summarize-func}

#' Given a dataframe, return a column-level summary of the mean and sd
#' of the numeric columns
#'
#' @param df A dataframe with numeric columns
#' @return A dataframe with three columns: name_col (column name), sd, and mean
summarize_data <- function(df) {
  funcs <- list(
    mean = ~ mean(.x),
    sd = ~ sd(.x),
    min = ~ min(.x),
    median = ~ median(.x),
    max = ~ max(.x)
  )
  mean_df <- summarize_func(df, "mean")
  sd_df <- summarize_func(df, "sd")
  min_df <- summarize_func(df, "min")
  median_df <- summarize_func(df, "median")
  max_df <- summarize_func(df, "max")
  # full_summary <- inner_join(mean_df, sd_df, by = "col_name")
  full_summary <- list(
    mean_df,
    sd_df,
    min_df,
    median_df,
    max_df
  ) %>% reduce(inner_join, by = "col_name")

  # full_summary <- inner_join(mean_df, sd_df, by = "col_name")
  return(full_summary)
}

#' Given a dataframe, return a column-level summary of the mean and sd
#' of the numeric columns
#'
#' @param df A dataframe with numeric columns
#' @return A dataframe with two columns: The original column name and the
#'   func summary of the original (numeric) column values
summarize_func <- function(df, func_name) {
  func <- get(func_name)
  namestring <- paste(func_name, "{.col}", sep = "")
  func_df <- df %>%
    summarize(across(where(is.numeric), func, .names = namestring)) %>%
    pivot_longer(
      cols = starts_with(func_name),
      names_to = "col_name",
      names_prefix = func_name,
      values_to = func_name,
      values_drop_na = FALSE
    )

  return(func_df)
}
```

```{r summary-table-with-income}
summarize_data(model_with_income_df) %>%
  column_to_rownames("col_name") %>%
  round(3) %>%
  kable("latex")
```

## Models

We will fit three types of model all with the same procedure, so we will make the process slightly easier as follows:

```{r model-procedure-func}
make_lms <- function(df) {
  ret_list <- vector(mode = "list")
  ret_list[["prop_MDE"]] <- MDE_lm(df)
  ret_list[["prop_suicidal_thoughts"]] <- suicidal_thoughts_lm(df)
  ret_list[["prop_alcohol_use_disorder"]] <- alcohol_use_disorder_lm(df)
  return(ret_list)
}
```

We can also extract the coefficient of interest as follows:

```{r tabulate-poverty-func}
extract_coef <- function(model, variable) {
  return(model$coefficients[[variable]])
}
extract_pval <- function(model, variable) {
  return(coef(summary(model))[variable, "Pr(>|t|)"])
}
format_confint <- function(model, variable) {
  # adjusted_model <- summary(model, test= adjusted(type = "holm"))
  # confidence <- confint(adjusted_model, variable, level = 0.05)
  confidence <- confint(model, variable, level = 0.95)
  return(
    paste("(", round(confidence[1], 3), ", ", round(confidence[2], 3), ")", sep = "")
  )
}
poverty_table <- function(model_list, adjust = FALSE, round = TRUE) {
  print(model_list[["prop_MDE"]]$coefficients[["prop_below_poverty_level"]])
  ret_table <- data.frame(
    c(
      "Major Depressive Episode (MDE)",
      "Suicidal Thoughts",
      "Alcohol Use Disorder"
    ),
    c(
      extract_coef(model_list[["prop_MDE"]], "prop_below_poverty_level"),
      extract_coef(model_list[["prop_suicidal_thoughts"]], "prop_below_poverty_level"),
      extract_coef(model_list[["prop_alcohol_use_disorder"]], "prop_below_poverty_level")
    ),
    c(
      extract_pval(model_list[["prop_MDE"]], "prop_below_poverty_level"),
      extract_pval(model_list[["prop_suicidal_thoughts"]], "prop_below_poverty_level"),
      extract_pval(model_list[["prop_alcohol_use_disorder"]], "prop_below_poverty_level")
    ),
    c(
      format_confint(model_list[["prop_MDE"]], "prop_below_poverty_level"),
      format_confint(model_list[["prop_suicidal_thoughts"]], "prop_below_poverty_level"),
      format_confint(model_list[["prop_alcohol_use_disorder"]], "prop_below_poverty_level")
    )
  )
  names(ret_table) <- c(
    "Response variable",
    "Coefficient",
    "$p$-value"
  )
  if (adjust) {
    ret_table[["$p$-value"]] <- p.adjust(ret_table[["$p$-value"]], "holm")
  }
  if (round) {
    ret_table[["Coefficient"]] <- round(ret_table[["Coefficient"]], 3)
    ret_table[["$p$-value"]] <- round(ret_table[["$p$-value"]], 3)
  }
  return(ret_table)
}
```

```{r table-primary}
poverty_table(make_lms(primary_model_df), adjust = TRUE) %>% kable("latex")
poverty_table(make_lms(primary_model_df), adjust = TRUE, round = FALSE) %>% kable("latex")
poverty_table(make_lms(model_with_income_df)) %>% kable("latex")
poverty_table(make_lms(model_with_state_df)) %>% kable("latex", escape = F)
poverty_table(make_lms(model_with_state_no_adjust_df)) %>% kable("latex", escape = F)
poverty_table(make_lms(model_with_state_no_adjust_df), round = F) %>% kable("latex", escape = F)
```

```{r tables-appendix}
confint_col <- function(model) {
  # adjusted_model <- summary(model, test= adjusted(type = "holm"))
  # confidence <- confint(adjusted_model, variable, level = 0.05)
  confidence <- confint(model, level = 0.95)
  return(
    paste("(", round(confidence[, 1], 3), ", ", round(confidence[, 2], 3), ")", sep = "")
  )
}
table_point_confint <- function(model) {
  coefs <- round(coef(model), 5)
  ret_table <- cbind(
    names(coefs),
    unname(coefs),
    confint_col(model)
  )
  colnames(ret_table) <- c("Variable", "Estimate", "95\\% CI")
  return(ret_table)
}
# table_point_confint(make_lms(primary_model_df)$prop_MDE) %>% kable("latex", escape=F, booktabs=T)
# table_point_confint(make_lms(primary_model_df)$prop_suicidal_thoughts) %>% kable("latex", escape=F, booktabs=T)
# table_point_confint(make_lms(primary_model_df)$prop_alcohol_use_disorder) %>% kable("latex", escape=F, booktabs=T)
```

### Primary models

```{r primary-models}
primary_model_list <- make_lms(primary_model_df)
```

```{r fitted-to-df}
fitted_primary_model_df <- primary_model_df
fitted_primary_model_df$prop_MDE_fitted <- fitted(primary_model_list[["prop_MDE"]])
fitted_primary_model_df$prop_suicidal_thoughts_fitted <- fitted(primary_model_list[["prop_suicidal_thoughts"]])
fitted_primary_model_df$prop_alcohol_use_disorder_fitted <- fitted(primary_model_list[["prop_alcohol_use_disorder"]])

map_fitted_model_df <- modified_full_df %>%
  full_join(fitted_primary_model_df)

# Maps of fitted values for primary model
map_and_save_var("./maps/prop_MDE_fitted.png", map_fitted_model_df, "prop_MDE_fitted", "Fitted values for primary MDE model", percent_flag = FALSE)
map_and_save_var("./maps/prop_suicidal_thoughts_fitted.png", map_fitted_model_df, "prop_suicidal_thoughts_fitted", "Fitted values for primary suicidal thoughts model", percent_flag = FALSE)
map_and_save_var("./maps/prop_alcohol_use_disorder_fitted.png", map_fitted_model_df, "prop_alcohol_use_disorder_fitted", "Fitted values for primary alcohol use disorder model", percent_flag = FALSE)

max_min_fitted_table <- cbind(
  c(
    "prop_MDE",
    "prop_suicidal_thoughts",
    "prop_alcohol_use_disorder"
  ),
  c(
    round(min(fitted_primary_model_df$prop_MDE_fitted), 3),
    round(min(fitted_primary_model_df$prop_suicidal_thoughts_fitted), 3),
    round(min(fitted_primary_model_df$prop_alcohol_use_disorder_fitted), 3)
  ),
  c(
    round(max(fitted_primary_model_df$prop_MDE_fitted), 3),
    round(max(fitted_primary_model_df$prop_suicidal_thoughts_fitted), 3),
    round(max(fitted_primary_model_df$prop_alcohol_use_disorder_fitted), 3)
  )
)
colnames(max_min_fitted_table) <- c("Variable", "Minimum fitted value", "Maximum fitted value")

max_min_fitted_table %>% kable("latex")
```

```{r resid-primary-plots}
resid_primary_model_df <- primary_model_df
resid_primary_model_df$prop_MDE_resid <- resid(primary_model_list[["prop_MDE"]])
resid_primary_model_df$prop_suicidal_thoughts_resid <- resid(primary_model_list[["prop_suicidal_thoughts"]])
resid_primary_model_df$prop_alcohol_use_disorder_resid <- resid(primary_model_list[["prop_alcohol_use_disorder"]])

map_resid_model_df <- modified_full_df %>%
  full_join(resid_primary_model_df)

map_and_save_var("./maps/prop_MDE_resid.png", map_resid_model_df, "prop_MDE_resid", "Residuals for primary MDE model", percent_flag = FALSE)

map_and_save_var("./maps/prop_alcohol_use_disorder_resid.png", map_resid_model_df, "prop_alcohol_use_disorder_resid", "Residuals for primary alcohol use disorder model", percent_flag = FALSE)

map_and_save_var("./maps/prop_suicidal_thoughts_resid.png", map_resid_model_df, "prop_suicidal_thoughts_resid", "Residuals for primary suicidal thoughts model", percent_flag = FALSE)
```

```{r resid-state-plots}
model_with_state_list <- make_lms(model_with_state_df)

resid_model_with_state_df <- model_with_state_df
resid_model_with_state_df$prop_MDE_resid <- resid(model_with_state_list[["prop_MDE"]])
resid_model_with_state_df$prop_suicidal_thoughts_resid <- resid(model_with_state_list[["prop_suicidal_thoughts"]])
resid_model_with_state_df$prop_alcohol_use_disorder_resid <- resid(model_with_state_list[["prop_alcohol_use_disorder"]])

map_resid_adj_model_df <- modified_full_df %>%
  full_join(resid_model_with_state_df)

map_and_save_var("./maps/prop_MDE_resid_adj.png", map_resid_adj_model_df, "prop_MDE_resid", "Residuals for state-adjusted alcohol use disorder model", percent_flag = FALSE)

map_and_save_var("./maps/prop_alcohol_use_disorder_resid_adj.png", map_resid_adj_model_df, "prop_alcohol_use_disorder_resid", "Residuals for state-adjusted alcohol use disorder model", percent_flag = FALSE)

map_and_save_var("./maps/prop_suicidal_thoughts_resid_adj.png", map_resid_adj_model_df, "prop_suicidal_thoughts_resid", "Residuals for state-adjusted suicidal thoughts model", percent_flag = FALSE)
```

### Secondary/Sensitivity Models

```{r generating-non-primary-models}
model_with_income_list <- make_lms(model_with_income_df)
model_with_state_list <- make_lms(model_with_state_df)
model_with_state_no_adjust_list <- make_lms(model_with_state_no_adjust_df)
```

### Variable correlations

```{r correlations}
cor(base_model_df$prop_MDE, base_model_df$prop_suicidal_thoughts)
cor(base_model_df$prop_MDE, base_model_df$prop_alcohol_use_disorder)
cor(base_model_df$prop_suicidal_thoughts, base_model_df$prop_alcohol_use_disorder)
```

# Presentation

We want a figure with the model coefficients and CIs as bars,
grouped by model group

```{r graph-pres}
make_df_from_model_group <- function(model_list) {
  estimate <- c()
  lower <- c()
  upper <- c()
  response <- c()
  for (dep_var_pair in list(
    list("prop_MDE", "Reported MDEs"),
    list("prop_suicidal_thoughts", "Reported suicidal thoughts"),
    list("prop_alcohol_use_disorder", "Reported alcohol abuse disorder")
  )) {
    dep_var <- dep_var_pair[[1]]
    dep_var_name <- dep_var_pair[[2]]
    model <- model_list[[dep_var]]
    coef_val <- coef(model)[["prop_below_poverty_level"]]
    confidence <- confint(model, parm = c("prop_below_poverty_level"), level = 0.95)
    response <- append(response, dep_var_name)
    estimate <- append(estimate, coef_val)
    lower <- append(lower, confidence[1])
    upper <- append(upper, confidence[2])
  }
  ret_df <- data.frame(response, estimate, lower, upper)
  return(ret_df)
}

overall_df_bar_graph <- data.frame()
for (pair in list(
  list(model_with_income_list, "Income-adjusted"),
  list(model_with_state_list, "State-adjusted (with additional)"),
  list(model_with_state_no_adjust_list, "Only state-adjusted"),
  list(primary_model_list, "Primary")
)) {
  model_list <- pair[[1]]
  model_group <- pair[[2]]
  print(model_list)
  df_with_model_group <- make_df_from_model_group(model_list)
  df_with_model_group$model_group <- model_group
  overall_df_bar_graph <- rbind(
    overall_df_bar_graph,
    df_with_model_group
  )
}
overall_df_bar_graph$response <- as.factor(overall_df_bar_graph$response)
overall_df_bar_graph$model_group <- as.factor(overall_df_bar_graph$model_group)

ggplot(data = overall_df_bar_graph) +
  geom_bar(aes(x = response, y = estimate, fill = response, group = response), stat = "identity", position = "dodge") +
  facet_wrap(. ~ model_group) +
  geom_errorbar(aes(x = response, ymin = lower, ymax = upper), width = 0.2) +
  theme_bw()
```
